---
title: 'Exercise 5: Functions and Loops'
author: "Dillon Laaker"
date: "11/5/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Thank you Veronica for assistance with some of the functions and with formulating the loops for part 2.

# Functions

1. Create a function for the mean, median, and standard deviation.


```{r message=FALSE, warning=FALSE}

national <- read.csv("national.csv")

library(tidyverse)

## Function for the mean ##

M <- function(x) {
  y <- (sum(x))/length(x)
  return(y)
  
}

## Function for the standard deviation ##

SD <- function(x) {
  y <- sqrt(sum((x - M(x))^2) / (length(x) - 1))
  return(y)
}

## Function for the median ## 

MD <- function(lst) {
  a <- length(lst)
  b <- sort(lst)
  y <- ifelse(a%%2==1,b[(a+1)/2], M(b[n/2+0:1]))
  return(y)
}

```

2. Create a function that finds the mean and excludes the lowest and highest value.


```{r message=FALSE, warning=FALSE}
M_exclude <- function(x) {
  x1 <- sort(x)
  y1 <- x1[-c(1, length(x1))]
  y2 <- M(x1)
  return(y2)
}

```

3. Apply the functions to the Christianity variables.


```{r message=FALSE, warning=FALSE}
## Mean ##

M(national$christianity_protestant) 
M(national$christianity_romancatholic)
M(national$christianity_easternorthodox)
M(national$christianity_anglican)
M(national$christianity_other)
M(national$christianity_all)

## Standard Deviation ##


SD(national$christianity_protestant) 
SD(national$christianity_romancatholic)
SD(national$christianity_easternorthodox)
SD(national$christianity_anglican)
SD(national$christianity_other)
SD(national$christianity_all)

## Median ## 


MD(national$christianity_protestant) 
MD(national$christianity_romancatholic)
MD(national$christianity_easternorthodox)
MD(national$christianity_anglican)
MD(national$christianity_other)
MD(national$christianity_all)

## Mean that excludes values ##

M_exclude(national$christianity_protestant) 
M_exclude(national$christianity_romancatholic)
M_exclude(national$christianity_easternorthodox)
M_exclude(national$christianity_anglican)
M_exclude(national$christianity_other)
M_exclude(national$christianity_all)


```

4. Write a function that lists all the unique years with more than 300,000 Christians in total.


```{r message=FALSE, warning=FALSE}

unique_years <- function(C) {
  y <- C %>% group_by(year, state) %>%
  summarise(C1 = christianity_all > 300000) %>%
  filter(C1 == T)
 return(y)  
}

y3 <- unique_years(national)

```

# Loops/apply

1. Write a loop to find how many variables there are per observation.

```{r message=FALSE, warning=FALSE}

LV <- rep(0, ncol(national))
for (i in 1:length(national)) {
  LV[i] <- ncol(national[i, ])
}

LV
```

2. Write a loop to find the mean number of Protestant Christians in each country (i.e., the state column). Then use an apply family function to do the same.


```{r message=FALSE, warning=FALSE}

MP <- data.frame(state = unique(national$state, prots = NA))
                 
for (i in 1:length(MP$state)) {
  state1 <- MP[i,1]
  y <- national %>% filter(state == state1) %>%
    summarise(mean = mean(christianity_protestant, na.rm = TRUE)) %>%
    pull(mean)
  MP[i, 2] <- y
  
}
```

3. Check the column type for each variable.

```{r message=FALSE, warning=FALSE}

sapply(national, class)

```
